### 2019-08-23 R/cfcs-example-reproducible-analysis.Rmd
added minimal reproducible example (CFCS)

---
title: "CFCS-Reproducible-Analysis-Example"
author: "Peikert and Brandmaier"
date: "8/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(lavaan)
library(digest)
```

## Loading data

```{r}
precomputed.digest <- "98c7e634dd7bb2ddf9d8f3972fff303365c55ac6"
temp <- tempfile()
download.file("http://openpsychometrics.org/_rawdata/CFCS.zip",destfile = temp)
dat <- read.table(unz(temp, "CFCS/data.csv"), header=TRUE)
unlink(temp)

if (precomputed.digest != digest::sha1(dat)) {
  error("Mismatch between original and current data file!")
}
```

## Fitting the model

```{r}
model <- paste0(" G =~ ", paste(paste0("Q",1:5),collapse="+"))

model.fit <- cfa(model, data=dat)

fitMeasures(model.fit)[c("cfi","rmsea")]

knitr::kable(lavaan::parameterEstimates(model.fit))
```