### 2019-10-28 R/rmarkdown.Rmd

---
title: "A simple RMarkdown Example"
author: "Aaron Peikert & Andreas M. Brandmaier"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: pdf_document
papersize: "a6"
geometry:
  - "landscape"
  - margin=0.5cm
header-includes:
   - \usepackage[center,dvips,frame]{crop}
---

```{r setup}
library("knitr")
library("papaja")
```

# Silly Heading

```{r t-test}
data("sleep")
result <- t.test(extra ~ group, data = sleep)
```

This is an example on students' sleep data taken from `help(t.test)`.

`r apa_print.htest(result)$full_result`

I can now assert that what I *believe* to be true
--- that there is a difference in means between the groups ---
is `r ifelse(result$p.value > .025, "**not**", "")` supported by data.