### 2019-10-29 abstract.Rmd
correct spelling

In this tutorial, we describe a principled workflow to ensure long-term reproducibility of R-based data analyses. The workflow leverages established tools and practices from software engineering. It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, APA-conformable dynamic report generation and full cross-platform and long-term computational reproducibility. The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.  While the research community increasingly recognizes dynamic document generation as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility. By ensuring full computational reproducibility together with version management, dependence management and literate programming,  the workflow increases scientific productivity by facilitating later reuse of code and data.