### 2019-11-11 README.Rmd
describe windows workaround close #43

---
output:
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

This is the accompanying GitHub repository to a work in progress paper by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982).

[![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0) [![Ask Me Anything !](https://img.shields.io/badge/Ask%20me-anything-1abc9c.svg)](https://github.com/aaronpeikert/reproducible-research/issues/new) ![Open Source Love](https://badges.frapsoft.com/os/v2/open-source.svg?v=103)

# Abstract

```{r abstract, child = 'abstract.Rmd'}
```

# Compile

## Usual Way

Requires: `Git`, `RStudio`, `pandoc`, `pandoc-citeproc` & `rmarkdown`.

Open RStudio -> File -> New Project -> Version Control -> Git

Insert:

```
https://github.com/aaronpeikert/reproducible-research.git
```

Open `manuscript.Rmd` click on `Knit`.

## Using Workflow

Does not require R or RStudio, but `make` & `docker`.

Execute in Terminal:

```{bash, eval=FALSE}
git clone https://github.com/aaronpeikert/reproducible-research.git
cd reproducible-research
make build
make all DOCKER=TRUE
```

**Note: Windows user need to hand edit the `Makefile` and set current_path to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`. We hope that future releases of Docker for Windows will not require that workaround.**

## Rebuild Everything

In case you experience some unexpected behavior with this workflow, you should check that you have the most recent version (`git pull`), rebuild the docker image (`make build`) and force the rebuild of all targets (`make -B DOCKER`).

```{bash, eval=FALSE}
git pull && make build && make -B DOCKER=TRUE
```

# Session Info

```{r}
sessioninfo::session_info()
```