### 2020-05-19 R/rmarkdown_fake.Rmd
update Fig 2 closes #39

---
title: "A simple R markdown example}"
author: "Aaron Peikert & Andreas M. Brandmaier"
output: pdf_document
---

### Silly Heading

```{r setup}
library("knitr")
library("papaja")
```

```{r t-test}
data("sleep")
result <- t.test(extra ~ group, data = sleep)
```

This is an example of students' sleep data taken from `help(t.test)`.

`r apa_print.htest(result)$full_result`

I can now assert that what I *believe* to be true
--- that there is a difference in means between the groups ---
is `r ifelse(result$p.value > .025, "**not**", "")` supported by the data.