### 2020-05-19 R/rmarkdown.Rmd

---
output: pdf_document
fontsize: 9pt
geometry:
  - "landscape"
  - margin=0.5cm
  - paperwidth=5cm
  - paperheight=10cm
header-includes:
   - \usepackage[center,dvips,frame]{crop}
---

### Dynamic Document Demonstration

```{r setup, echo=FALSE}
library("knitr")
library("papaja")
```

This is a simple analysis of the `sleep` dataset (Student, 1908) taken from `help(t.test)`.

```{r t-test}
data("sleep")
result <- t.test(extra ~ group, data = sleep, paired = TRUE)
```

The difference in means of hours slept between the groups 
was `r ifelse(result$p.value >= .05, "**not**", "")` 
significantly different from zero (`r apa_print.htest(result)$full_result`).